### 2017-06-18 paper.tex
Second round of excess word removal and grammar cleaning

The first requirement for building lessons together is
to know who they are being built for.
``Archaeology students'' is far too vague:
are you and your collaborators thinking of
first-year students who need an introduction to the field,
graduate students who intend to specialize in the sub-discipline which is the lesson's focus,
or someone in between?
Prerequisite knowledge,
equipment or software required,
how much time learners will actually have:
if different contributors believe different things about these,
they will find it difficult or impossible to work together.

@@

Lessons don't maintain themselves.
For technical lessons, software versions and dependencies are constantly changing,
while for more traditional academic lessons the literature is advancing at an ever increasing pace.
In either case, what was cutting edge in 2017 may be dated
and less useful in 2018.
This needs to be clear to contributors,
so they understand that this lesson is just
the starting point.
Sustainability needs to be front of mind.

The focus accordingly needs to be on creating a community.
Authors cannot be expected to maintain continual vigilance on a lesson,
but this is necessary if one expects continual use of a lesson!
By working online, creating opportunities for collaboration
and contribution, many eyes can keep the lesson usable.
As we note below,
this only works if you make space for
little contributions just as you do for larger ones.

@@

Experience shows that even a brief introduction to a handful of key practices
can help collaborative lesson development in at least three ways.
If people have a shared understanding of how lessons should be developed,
it is easier for them to work together.
Less obviously,
if people have a shared model of how lessons are going to be \emph{used},
they are more likely to try to build the same kind of material.
Finally,
teaching people how to teach is a great way to introduce them to each other and build community.

@@

Making the process for contributing to a lesson explicit is
the key to actually getting contributions.
New contributors require a straightforward and transparent introduction
to understand the process of tweaking and adapting lessons.
Licensing, code of conduct, governance, and contribution procedure
all need to be explicit rather than implicit
to lower the social barriers to contribution.

@@

The best way to choose tools for managing lessons
is to look for those that provide a gentle on-ramp for new contributors.
Ask potential contributors what they are comfortable with
rather than requiring \emph{them} to come to \emph{you}.
Remember also that contributing to a lesson is probably not their top priority,
and look for ways to reduce their cognitive load.
For example,
threaded discussion on particular topics (e.g., using GitHub issues)
can help increase the signal-to-noise ratio
by reducing long reply-all email exchanges. 
A number of open frameworks are available to facilitate  
development of new lessons, such as R-based
learnr (\url{https://rstudio.github.io/learnr}) or 
GitHub-based Morea (\url{https://morea-framework.github.io}) and 
DataCamp (\url{https://www.datacamp.com/teach/documentation}).

@@

Micro-scale feedback can be gathered by an instructor while teaching a particular lesson.
Learners might provide feedback on things like typographical errors,
clarity/ease of quiz questions and/or the order in which topics are presented,
which the instructor can enter into a work-tracking system (e.g., GitHub issues) at the end of class.
As well as encouraging direct verbal feedback,
it's a good idea to provide learners with a means to provide feedback anonymously during class
(e.g., on small pieces of paper like sticky notes).

Pre- and post-class surveys can be used to discover larger macro-scale issues.
These issues often relate to the fact that it can be difficult for lesson developers
to fully understand the frame of reference of their audience.
For instance, a lesson might inadvertently assume prior knowledge that learners don't have,
which is information that can be collected in a survey.
If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
This allows people time to reflect,
meaning they are more likely to give accurate feedback on what they learned
rather than how entertained they were.

@@

Don't create a new lesson if there is an existing one that you can use or contribute to.
Just as a scholar would not write a paper without a literature review,
the same holds for lessons.
Before writing that introduction to the Bash command line,
for example,
do a search:
has anybody else written it?
Is it complementary to your goals?
Could it be tweaked or modified to meet your own goals?
Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?

@@

Doing this is a lot of work,
and maintaining it is even more so,
which makes building community around lessons (Rule~2) all the more important.
In particular,
it is vital to engage the learners as equal participants in that community:
they should both be able to propose updates, corrections, and additions to lessons,
and know that they are encouraged to do so (Rule~5).

@@

If there are complementary ways to explain something,
or points of view that can cohabit respectfully,
it may be possible to present them side by side.
There are actually good pedagogical reasons to do this
even if contributors \emph{don't} disagree:
weighing alternatives fosters higher-order thinking,
and as Caulfield has written,
the most popular online question and answer sites
are successful in part because they present a chorus of explanations
geared at different levels and needs
rather than a single one \cite{choral-explanations}.

But sometimes choices must be made.
The open source software community has wrestled with these issues for three decades,
and has evolved some best practices to handle them
\cite{producing-oss}.
As discussed in Rule~5,
the first step is to have a clear governance structure and a clear, permissive license.
Minor disagreements should be discussed openly and respectfully.
If they cannot be resolved---i.e., if they turn out not to be so minor after all---then
contributors should split off and evolve the lesson in the way they see best.
(This is one of the reasons to have a permissive license.)

>>>>

The first requirement for building lessons together is
to know who they are being built for.
``Archaeology students'' is far too vague:
are you and your collaborators thinking of
first-year students who need an introduction to the field,
graduate students who intend to specialize in the sub-discipline which is the lesson's focus,
or someone in between?
Prerequisite knowledge,
equipment or software required,
how much time learners will have:
if different contributors believe different things about these,
they will find it difficult or impossible to work together.

@@

Lessons do not maintain themselves.
For technical lessons, software versions and dependencies are constantly changing,
while for more traditional academic lessons the literature is advancing at an ever increasing pace.
In either case, what was cutting edge in 2017 may be dated
and less useful in 2018.
This needs to be clear to contributors,
so they understand that this lesson is just
the starting point.
Sustainability needs to be front of mind.

The focus accordingly needs to be on creating a community.
Authors cannot be expected to maintain continual vigilance on a lesson,
but this is necessary if one expects continual use of a lesson!
By working online, creating opportunities for collaboration
and contribution, many eyes can keep the lesson usable.
As we describe in Rule~5,
this only works if you make space for
little contributions just as you do for larger ones.

@@

Experience shows that even a brief introduction to a handful of key practices
can help collaborative lesson development in at least three ways.
If people have a shared understanding of how lessons should be developed,
it is easier for them to work together.
Less obviously,
if people have a shared model of how lessons will be \emph{used},
they are more likely to try to build the same kind of material.
Finally,
teaching people how to teach is a great way to introduce them to each other and build community.

@@

Making the process for contributing to a lesson explicit is
the key to receiving contributions.
New contributors require a straightforward and transparent introduction
to understand the process of tweaking and adapting lessons.
Licensing, code of conduct, governance, and contribution procedure
all need to be explicit rather than implicit
to lower the social barriers to contribution.

@@

The best way to choose tools for managing lessons
is to look for those that provide a gentle on-ramp for new contributors.
Ask potential contributors what they are comfortable with
rather than requiring \emph{them} to come to \emph{you}.
Remember also that contributing to a lesson is probably not their top priority,
and look for ways to reduce their cognitive load.
For example,
threaded discussion on particular topics (e.g., using GitHub issues)
can help increase the signal-to-noise ratio
by reducing long reply-all email exchanges.
Several open frameworks are available to facilitate
development of new lessons, such as R-based
learnr (\url{https://rstudio.github.io/learnr}) or
GitHub-based Morea (\url{https://morea-framework.github.io}) and
DataCamp (\url{https://www.datacamp.com/teach/documentation}).

@@

Micro-scale feedback can be gathered by an instructor while teaching a particular lesson.
Learners might provide feedback on things like typographical errors,
clarity/ease of quiz questions or the order in which topics are presented,
which the instructor can enter into a work-tracking system (e.g., GitHub issues) at the end of class.
As well as encouraging direct verbal feedback,
it's a good idea to provide learners with a means to provide feedback anonymously during class
(e.g., on small pieces of paper like sticky notes).

Pre- and post-class surveys can be used to discover larger macro-scale issues.
These issues often relate to the fact that it can be difficult for lesson developers
to fully understand the frame of reference of their audience.
For instance, a lesson might inadvertently assume prior knowledge that learners do not have,
which is information that can be collected in a survey.
If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
This delay allows people time to reflect,
meaning they are more likely to give accurate feedback on what they learned
rather than how entertained they were.

@@

Do not create a new lesson if there is an existing one that you can use or contribute to.
Just as a scholar would not write a paper without a literature review,
the same holds for lessons.
Before writing that introduction to the Bash command line,
for example,
do a search:
has anybody else written it?
Is it complementary to your goals?
Could it be tweaked or modified to meet your own goals?
Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?

@@

Doing this is substantial work,
and maintaining it is even more so,
which makes building community around lessons (Rule~2) all the more important.
In particular,
it is vital to engage the learners as equal participants in that community:
they should both be able to propose updates, corrections, and additions to lessons,
and know that they are encouraged to do so (Rule~5).

@@

If there are complementary ways to explain something,
or points of view that can cohabit respectfully,
it may be possible to present them side by side.
There are good pedagogical reasons to do this
even if contributors \emph{do not} disagree:
weighing alternatives fosters higher-order thinking,
and as Caulfield has written,
the most popular online question and answer sites
are successful in part because they present a chorus of explanations
geared at different levels and needs
rather than a single one \cite{choral-explanations}.

But sometimes choices must be made.
The open source software community has wrestled with these issues for three decades,
and has evolved some best practices to address them
\cite{producing-oss}.
As discussed in Rule~5,
the first step is to have a clear governance structure and a clear, permissive license.
Minor disagreements should be discussed openly and respectfully.
If they cannot be resolved---i.e., if they turn out not to be so minor after all---then
contributors should split off and evolve the lesson in the way they see best.
(This is one of the reasons to have a permissive license.)
